Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-08-14 grove-paper.tex
Work on Figures for "Addition" and "Change" section

@@

TODO: Screenshot

\begin{figure}
\centering
TODO: UI

\includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
\caption{A simple tree with a hole}
\label{fig:A simple tree with a hole}
\end{figure}

\begin{figure}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[tree layout] {
 root[as={\textbullet}] -> {
  +[as={$\texttt{+}_1$}] -> [edge label={left}] {
   22[as={$\texttt{num(22)}_2$}],
   {}
  }
 }
};
\end{tikzpicture}
\caption{A simple tree with a hole (Alternate figure)}
\label{fig:A simple tree with a hole (Alternate figure)}
\end{figure}

The editor state for each replica in Grove consists of an expression with holes.
For example, Figure~TODO:1(a) shows the simple arithmetic expression \texttt{2 + \hole}, which
has a hole on the right side of the plus operator.

@@

@@

TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)

TODO: remove cursor from Fig 1 and Fig 2

@@

Continuity of examples: Add -> Change -> Wrap -> Move

@@

TODO: num(22) -> var(x)

@@

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small}}
}}
\tikzset{graphs/root style/.style={
 draw=none,
 as={\textbullet}
}}
%\tikzset{graphs/every edge/.style={}
\begin{figure}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_2$}],
   y [> edge label={right}, > edge label'={3}, as={$\texttt{var(y)}_3$}]
  }
 }
};
\end{tikzpicture}
\caption{Addition}
\label{fig:Addition}
\end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
  }
 }
};
\path (1cm,0cm) graph[graph style] {
 y [as={$\texttt{var(y)}_3$}]
};
\end{tikzpicture}
\caption{TODO}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
   z [> edge label={right}, as={$\texttt{var(z)}_4$}]
  }
 }
};
\path (1cm,0cm) graph[graph style] {
 y [as={$\texttt{var(y)}_3$}]
};
\end{tikzpicture}
\caption{TODO}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

TODO: 2 -> 22

We model edits to the code as sequences of deletions and additions of edges to this graph.
For example, consider the first graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + 2}.
Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
becomes \texttt{1 + 3}.
We model this by first deleting edge~7.
This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + \hole}.
Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
Vertex~9 is just not reachable from the root.

>>>>

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small}}
}}
\tikzset{graphs/root style/.style={
 draw=none,
 as={\textbullet}
}}
%\tikzset{graphs/every edge/.style={}

@@

% \begin{figure}
% \centering
% TODO: UI
% \includegraphics[scale=0.75]{graphviz/simple-tree-with-hole/1.dot}
% \caption{A simple tree with a hole}
% \label{fig:A simple tree with a hole}
% \end{figure}

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:A simple tree with a hole-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> [] {
   22[> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_3$}],
   {}
  }
 }
};
\end{tikzpicture}
\caption{Graph representation}
\label{fig:A simple tree with a hole-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{A simple tree with a hole}
\label{fig:A simple tree with a hole}
\end{figure}

The editor state for each replica in Grove consists of an expression with holes.
For example, Figure~\ref{fig:A simple tree with a hole} shows the simple arithmetic expression \texttt{2 + \hole}, which
has a hole on the right side of the plus operator.

@@

For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

@@

% TODO: change num(2) in Fig1 to num(1) so Fig1 matches Fig2 (move parts of Fig 2 into Fig 1?)

% TODO: remove cursor from Fig 1 and Fig 2

@@

@@

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:Addition-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, > edge label'={2}, as={$\texttt{var(x)}_2$}],
   y [> edge label={right}, > edge label'={3}, as={$\texttt{var(y)}_3$}]
  }
 }
};
\end{tikzpicture}
\caption{Graph representation}
\label{fig:Addition-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Addition}
\label{fig:Addition}
\end{figure}

We model edits to the code as sequences of additions and deletions of edges to this graph.
For example, consider the graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + 2}.
Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
becomes \texttt{1 + 3}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 y [as={$\texttt{var(y)}_3$}]
};
\end{tikzpicture}
\caption{TODO}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  +[as={$\texttt{+}_1$}] -> {
   x [> edge label={left}, as={$\texttt{var(x)}_2$}],
   z [> edge label={right}, as={$\texttt{var(z)}_4$}]
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 y [as={$\texttt{var(y)}_3$}]
};
\end{tikzpicture}
\caption{TODO}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

We model this by first deleting edge~7.
This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + \hole}.
Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
Vertex~9 is just not reachable from the root.
